Value Ordering, Discrepancies, and Restarts in Subgraph Solvers: \\ Some Half-Baked Mucking Around

### 2017-11-09 paper/paper.tex
All by family

We could try to introduce tiebreaking mechanisms to further refine this heuristic. However, given
the extreme cost of making an incorrect branching choice early on in search, we believe it is more
fruitful to investigate alternatives to simple backtracking (depth-first) search. This paper
evaluates two such approaches: discrepancy search (which perform poorly), and restarts (which are
extremely beneficial). To make restarts work, we introduce a new value-ordering heuristic, which is
nearly the same as ``highest degree first'' except with small amounts of heavily biased randomness.
The effects of the changes we propose are shown in \cref{figure:scatter}: we can make a
state-of-the-art algorithm perform much better on a large number of satisfiable instances, without
substantially degrading its performance on unsatisfiable instances.

@@

\begin{figure}[p]
    \centering
    \includegraphics*{gen-graph-scatter.pdf}

    \caption{Instance by instance comparison of the basic algorithm with the degree heuristic,
    versus degree-biased + restarts.}
    \label{figure:scatter}
\end{figure}

@@

In the more detailed view in \cref{figure:scatter}, comparing the basic algorithm with the degree
heuristic to the degree-biased algorithm with restarts, all of the unsatisfiable instances
are on the $x-y$ diagonal, showing that their performance is unchanged. On the other hand, there are
?? large numbers of satisfiable instances well below the diagonal line, indicating large speedups.
Better yet, there are only ?? satisfiable instances that are more than a factor of ?? times worse.

@@

\begin{figure}[tb]
    \centering
    \includegraphics*{gen-graph-scatter-by-family.pdf}

    \caption{Instance by instance comparison of the basic algorithm with the degree heuristic,
    versus degree-biased + restarts, showing only satisfiable instances, on a family by family
    basis.}
    \label{figure:scatter-by-family}
\end{figure}

?? do a more detailed breakdown by family, \cref{figure:scatter-by-family}

>>>>

We could try to introduce tiebreaking mechanisms to further refine this heuristic. However, given
the extreme cost of making an incorrect branching choice early on in search, we believe it is more
fruitful to investigate alternatives to simple backtracking (depth-first) search. This paper
evaluates two such approaches: discrepancy search (which perform poorly), and restarts (which are
extremely beneficial). To make restarts work, we introduce a new value-ordering heuristic, which is
nearly the same as ``highest degree first'' except with small amounts of heavily biased randomness.
The effects of the changes we propose are shown in \cref{figure:scatter-by-family}: we can make a
state-of-the-art algorithm perform much better on a large number of satisfiable instances, without
substantially degrading its performance on unsatisfiable instances.

@@

\begin{figure}[p]
    \centering
    \includegraphics*{gen-graph-scatter-by-family.pdf}

    \caption{Instance by instance comparison of the basic algorithm with the degree heuristic,
    versus degree-biased + restarts.}
    \label{figure:scatter-by-family}
\end{figure}

@@

In the more detailed view in \cref{figure:scatter-by-family}, comparing the basic algorithm with the
degree heuristic to the degree-biased algorithm with restarts, all of the unsatisfiable instances
are very close to the $x-y$ diagonal, showing that their performance is nearly unchanged. On the
other hand, there are ?? large numbers of satisfiable instances well below the diagonal line,
indicating large speedups.  Better yet, there are only ?? satisfiable instances that are more than a
factor of ?? times worse.

?? By family

@@
